---
title: 'class 11: scraping multiple webpages'
author: "zina dupal"
date: "2026-09-30"
output: 
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    code_download: true
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Learning obj for today

- Scrape multiple webpages using `for` loops; 

- Scrape multiple webpages using the `purrr::map()` type functions, which would also make you write your own functions in R. 

## Extending the course bulletin scrape for multiple departments

We will scrape the following pages:
[ISA](https://bulletin.miamioh.edu/courses-instruction/isa/) and [MKT](https://bulletin.miamioh.edu/courses-instruction/mkt/). From the login shown in class, you can do that for all FSB departments. 

If you are doing it for just a few departments, you can just repeat your code a few times; however, this is bad practice so we will not do this in class. Software engineers tend to refer this as: **DRY** (don't repeat yourself).

```{r bulletin}
dept_abbr = c('acc', 'bus', 'isa', 'mkt') # extending the example to 4 depts

# line below illustrates how the url can be constructed by pasting 0, i.e., no space, the base_url, dep_abbreviation and the forward slash

courses_df = data.frame() # initialize the container
for (dept_name in 1:length(dept_abbr)) {
  cat(rep("-",25))
  print(dept_name)
  dept_url = paste0(
    "https://bulletin.miamioh.edu/courses-instruction/", dept_abbr[dept_name], "/")
  
  # for the titles, use the selector gadget and for the desc, using the inspector (to show the alternative CSS selectors for what we did on Mon)
  webpage = rvest::read_html(dept_url)
  
  webpage |> rvest::html_elements("strong") |> rvest::html_text2() -> titles
  webpage |> 
    rvest::html_elements("#sc_sccoursedescs > div > p.courseblockdesc") |>
    rvest::html_text2() -> descs
  
  df = data.frame(title = titles, desc = descs)
  df$title[1] |> print() # showing only the first course title
  courses_df = rbind(courses_df, df)
  cat(rep("-",25))
}
```

## Re-examining the prior example with the `purrr` package

My preferred approach with the purrr fn is to first create my own function and then use it within the package. 

```{r}
scrape_mu_bulletin = function(dept_url) {

  dept_url = paste0(
    "https://bulletin.miamioh.edu/courses-instruction/", dept_abbr[dept_name], "/")
  
  # for the titles, use the selector gadget and for the desc, using the inspector (to show the alternative CSS selectors for what we did on Mon)
  webpage = rvest::read_html(dept_url)
  
  webpage |> rvest::html_elements("strong") |> rvest::html_text2() -> titles
  webpage |> 
    rvest::html_elements("#sc_sccoursedescs > div > p.courseblockdesc") |>
    rvest::html_text2() -> descs
  
  df = data.frame(title = titles, desc = descs)
  
  return(df)
}

test = scrape_mu_bulletin("https://bulletin.miamioh.edu/courses-instruction/bus/")

dept_abbr = c('acc', 'bus', 'isa', 'mkt')
my_urls = paste0(
  "https://bulletin.miamioh.edu/courses-instruction/", # single chr string
  dept_abbr, # vector of 4 obs
  "/" # single chr string
)

# the map fns take 1 or more inputs (depending on the fns you are using)
# in this example, we only have 1 input for our function
# we used map_() instead of map2_() [2 is if you have two inputs], p_map() [for 3 or more inputs]
# .x = the vectore you are iterating over
courses_df_from_purrr = purrr::map_df(.x = my_urls, .f = scrape_mu_bulletin)

head(courses_df_from_purrr)
```

