---
title: "hw 04"
author: "zina dupal"
date: "2026-09-13"
output:
  html_document:
    toc: true
    toc_float: true
    numbered_sections: true
    code_folding: show
    code_download: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Assignment 04: Subsetting and Import

```{r}
con = DBI::dbConnect(RSQLite::SQLite(), "../data/midwest_airbnb.db")
listings = DBI::dbReadTable(con, "listings")
```

## A1: How many listings (rows) does `listings` have?
```{r A1}
nrow(listings)
```

## A2: How many of the listings are in Columbus? The `city` column is `"Chicago"`, `"Columbus"`, or `"Twin Cities"`.
```{r A2}
sum(listings$city == "Columbus")
```

## A3: Which of these is the name of the most expensive listing in Chicago (highest price, in US dollars per night)? All four are real Chicago listings from the data.
```{r A3}
chicago = listings[listings$city == "Chicago", ]
chicago$name[which.max(chicago$price)]
```

## A4: How many distinct values does the `neighbourhood` column contain among the Chicago listings? 
```{r A4}
length(unique(listings$neighbourhood[listings$city == "Chicago"]))
```

## A5: How many listings are run by a superhost? The `host_is_superhost` column is `"t"` for a superhost, `"f"` otherwise, and `NA` for a few hosts, so tell `sum()` to skip the `NA`s.
```{r A5}
sum(listings$host_is_superhost == "t", na.rm = TRUE)
```

## A6: How many listings cost more than $1,000 per night? The `price` column is numeric, in US dollars per night.
```{r A6}
sum(listings$price > 1000)
```

## B1: Which statement is correct about reading a CSV file from a URL and from a file on disk? 
#### **readr::read_csv() works for both: pass a URL or a file path as the same first argument**

## B2: In class, the BLS occupation workbook had to be downloaded in a browser before R could read it. Why? 
#### **read_excel() reads only files on disk, and the BLS site blocks scripted downloads** 

## B3: Your project has the folders `markdowns/` and `data/` side by side. When you knit `markdowns/class05.Rmd`, which path reaches `data/occupation.xlsx`? 
#### **../data/occupation.xlsx**

## B4: What does `DBI::dbReadTable(con, "listings")` return? 
#### **A data frame with one row per record in the table**


