---
title: "Airbnb Analysis"
author: "zina dupal"
date: "2026-09-24"
output: 
  html_document:
    toc: true
    toc_float: true
    numbered_sections: true
    code_folding: show
    code_download: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r data}
con = DBI::dbConnect(
  RSQLite::SQLite(),
  "../data/midwest_airbnb_extended.db"
)

listings = DBI::dbReadTable(con, "listings") |>
  tibble::as_tibble()

availability = DBI::dbReadTable(con, "availability_monthly") |>
  tibble::as_tibble()

reviews = DBI::dbReadTable(con, "reviews_monthly") |>
  tibble::as_tibble()

DBI::dbDisconnect(con)
```

## Task 1: Who owns the market

```{r task1}
host_counts = listings |>
  dplyr::count(host_id, name = "host_listings")

listings_hosts = listings |>
  dplyr::left_join(host_counts, by = "host_id")

host_counts |>
  dplyr::filter(host_listings >= 10) |>
  dplyr::summarise(A1 = dplyr::n())

listings_hosts |>
  dplyr::filter(city == "Chicago") |>
  dplyr::summarise(
    A2 = mean(host_listings >= 10) * 100
  )
```

## Task 2: Value for a family of four

```{r task2}
listings |>
  dplyr::filter(
    room_type == "Entire home/apt",
    accommodates >= 4,
    number_of_reviews >= 10,
    review_scores_rating >= 4.8
  ) |>
  dplyr::mutate(
    price_per_person = price / accommodates
  ) |>
  dplyr::group_by(city, neighbourhood) |>
  dplyr::summarise(
    median_price_per_person = median(price_per_person, na.rm = TRUE),
    n = dplyr::n()
  ) |>
  dplyr::filter(
    n >= 20,
    city == "Columbus"
  ) |>
  dplyr::arrange(median_price_per_person)
```

## Task 3: What earns

```{r task3}
listings |>
  dplyr::group_by(city, room_type) |>
  dplyr::summarise(
    dplyr::across(
      c(price, reviews_per_month, estimated_revenue_l365d),
      ~ median(.x, na.rm = TRUE)
    )
  ) |>
  dplyr::filter(
    city == "Twin Cities",
    room_type == "Entire home/apt"
  )
```

## Task 4: Concentration

```{r task4}
top_listings = listings |>
  dplyr::group_by(city) |>
  dplyr::arrange(
    dplyr::desc(estimated_revenue_l365d),
    .by_group = TRUE
  ) |>
  dplyr::mutate(
    rank = dplyr::row_number(),
    cutoff = ceiling(0.10 * dplyr::n())
  ) |>
  dplyr::filter(rank <= cutoff)

top_revenue = top_listings |>
  dplyr::filter(city == "Columbus") |>
  dplyr::summarise(
    top_revenue = sum(estimated_revenue_l365d, na.rm = TRUE)
  )

total_revenue = listings |>
  dplyr::filter(city == "Columbus") |>
  dplyr::summarise(
    total_revenue = sum(estimated_revenue_l365d, na.rm = TRUE)
  )

top_revenue$top_revenue / total_revenue$total_revenue * 100
```

## Task 5: The year ahead

```{r task5}
october = availability |>
  dplyr::filter(month == "2026-10")

october_listings = listings |>
  dplyr::filter(room_type == "Entire home/apt") |>
  dplyr::inner_join(october, by = "id")

october_listings |>
  dplyr::group_by(city) |>
  dplyr::summarise(
    share_mostly_taken =
      mean(nights_available < 0.5 * nights_in_month) * 100
  ) |>
  dplyr::arrange(share_mostly_taken)
```

## Task 6: Demand by season

```{r task6}
monthly_reviews = reviews |>
  dplyr::filter(
    month >= "2025-07",
    month <= "2026-06"
  ) |>
  dplyr::inner_join(listings, by = "id") |>
  dplyr::group_by(city, month) |>
  dplyr::summarise(
    total_reviews = sum(reviews),
    .groups = "drop"
  )

twin_cities = monthly_reviews |>
  dplyr::filter(city == "Twin Cities") |>
  dplyr::arrange(dplyr::desc(total_reviews))

twin_cities |>
  dplyr::slice_max(total_reviews, n = 1)

twin_cities |>
  dplyr::summarise(
    A9 = max(total_reviews) / min(total_reviews)
  )
```

## Task 7: The Superhost premium

```{r task7}
superhost_summary = listings |>
  dplyr::filter(
    room_type == "Entire home/apt",
    host_is_superhost %in% c("t", "f")
  ) |>
  dplyr::group_by(city, host_is_superhost) |>
  dplyr::summarise(
    median_price = median(price, na.rm = TRUE),
    median_rating = median(review_scores_rating, na.rm = TRUE),
    .groups = "drop"
  )

superhost_summary

superhost_summary |>
  dplyr::group_by(city) |>
  dplyr::summarise(
    premium =
      median_price[host_is_superhost == "t"] -
      median_price[host_is_superhost == "f"]
  )
```

## Task 8: What the join lost

```{r task8}
unmatched = listings |>
  dplyr::anti_join(availability, by = "id")

unmatched |>
  dplyr::summarise(
    A10 = dplyr::n()
  )

unmatched |>
  dplyr::count(city) |>
  dplyr::arrange(dplyr::desc(n))
```

## B1, the badge.

> Chicago: $43.80; Columbus: $40.50; Twin Cities: -$1.38. In the Twin Cities, Superhosts charge a median price about $1.38 lower than non-Superhosts, so the Superhost badge is not associated with a higher median price in the Twin Cities. Before determining if the badge is not worth it, I would also compare the median review rating for Twin Cities Superhosts and non-Superhosts.

## B2, the cap.

```{r b2}
# Task 1 share
host_share = listings_hosts |>
  dplyr::group_by(city) |>
  dplyr::summarise(
    host_10_share = mean(host_listings >= 10) * 100
  )

host_share

# Task 4 share
revenue_share = listings |>
  dplyr::group_by(city) |>
  dplyr::arrange(
    dplyr::desc(estimated_revenue_l365d),
    .by_group = TRUE
  ) |>
  dplyr::mutate(
    rank = dplyr::row_number(),
    cutoff = ceiling(0.10 * dplyr::n())
  ) |>
  dplyr::summarise(
    top_10_rev_share =
      sum(
        estimated_revenue_l365d[rank <= cutoff],
        na.rm = TRUE
      ) /
      sum(estimated_revenue_l365d, na.rm = TRUE) * 100
  )

revenue_share
```

> Columbus would be affected most because hosts with 10 or more listings control 37.1% of its listings, compared with 31.9% in Chicago and 17.2% in the Twin Cities; its top 10% of listings also earn 36.1% of estimated revenue, compared with 40.7% in Chicago and 40.3% in the Twin Cities. However, this data cannot tell the council whether one host ID always represents one business.


